\documentclass[11pt]{article}

\usepackage{acl}
\usepackage{times}
\usepackage{latexsym}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{inconsolata}
\usepackage{graphicx}
\usepackage{booktabs}

\title{Global-Context Pointwise Reranking under Limited Compute: An Empirical Study with Qwen2.5 on SciFact}

\author{
  Alwie Attar Elfandra \\
  \texttt{2306241726}
  \And
  Daffa Naufal Rahadian \\
  \texttt{2306213003}
  \And
  Edmond Christian \\
  \texttt{2306208363}
}

\begin{document}
\maketitle

\begin{abstract}
Large Language Models (LLMs) have shown remarkable capabilities as zero-shot rerankers in information retrieval, yet their heavy computational footprints often make them impractical for resource-constrained environments. This work evaluates a resource-optimized implementation of global-context pointwise reranking inspired by GCCP, a method that enhances independent pointwise scoring by introducing collection-wide context. We conduct comprehensive completed experiments on the SciFact dataset using ultra-lightweight Small Language Models (SLMs), specifically \texttt{Qwen2.5-1.5B-Instruct} and \texttt{Qwen2.5-3B-Instruct}, deployed under 4-bit NormalFloat4 (NF4) quantization. Additionally, we integrate an advanced spectral sentence-anchor extension utilizing graph Laplacian decomposition to construct highly coherent global context. To circumvent Out-of-Memory (OOM) errors and latency bottlenecks on a single commodity GPU, we leverage a customized left-padding batching inference strategy. Our empirical results demonstrate that while standalone pointwise SLM ranking substantially lags behind BM25, the integration of global context yields significant performance leaps. Notably, the \texttt{Qwen2.5-3B-Instruct} backbone augmented with our Spectral GCCP and PAGC framework achieves a breakthrough nDCG@10 of \textbf{0.6624}, successfully outperforming the highly competitive BM25 baseline (0.6519) while maintaining a safe and optimized execution profile.
\end{abstract}

\section{Introduction}

Reranking is a central component in modern information retrieval systems. A first-stage retriever such as BM25 retrieves a candidate set efficiently, and a second-stage model reorders those candidates with a richer relevance function. Recent work has shown that large language models (LLMs) can be used as zero-shot rerankers by prompting the model to judge whether a passage is relevant to a query. Pointwise reranking is especially attractive because it scores each query-document pair independently, making the number of model calls linear in the number of candidates.

The limitation of pointwise reranking is that every document is evaluated in isolation. In a difficult candidate set, a document may look relevant when read alone but less useful when compared with the recurring evidence across other candidates. GCCP addresses this limitation by constructing global context from the candidate set and then using post-aggregated contrastive scoring \citep{long2025gccp}. While the original framework was evaluated on high-parameter models across multi-GPU clusters, its viability using ultra-small instruction-tuned models in severely constrained computing environments remains unexplored.

This project investigates that question in a constrained experimental setting. Instead of claiming a full reproduction of the original LLM-based GCCP pipeline, we implement a modular retrieval framework that can run within restricted local hardware and exposes an operational LLM scoring interface. The default configurations evaluate the performance of \texttt{Qwen2.5-1.5B-Instruct} and \texttt{Qwen2.5-3B-Instruct} under 4-bit NF4 quantization on the BEIR SciFact dataset. 

Our main research question is: can GCCP-style global context improve pointwise reranking under limited compute? We also add a small methodological extension: a spectral sentence-anchor strategy that uses a TF-IDF sentence graph and the Fiedler vector of the normalized Laplacian to select a coherent anchor from the candidate set. This extension is compared with a simpler sentence-overlap anchor.

\section{Related Work}

BM25 is a classical and highly competitive lexical retrieval method, particularly for collections where exact terminology remains informative \citep{robertson2009probabilistic}. In many retrieval pipelines, BM25 is used as an efficient first-stage retriever before neural reranking. The BEIR benchmark was introduced to evaluate retrieval systems across heterogeneous domains, including scientific fact verification through SciFact \citep{thakur2021beir}.

LLM-based reranking has become popular because instruction-tuned models can judge relevance from natural-language prompts without task-specific training. Pointwise prompting asks the model whether a document is relevant to the query. Pairwise and listwise methods give the model multiple documents at once, but they are often more expensive or harder to scale. GCCP attempts to preserve the simplicity of pointwise reranking while injecting global information from the candidate set \citep{long2025gccp}. Our work follows this motivation but focuses on a small, reproducible implementation suitable for resource-limited machines.

\section{Methodology}

\paragraph{First-stage retrieval.}
We use BM25 to retrieve the top 100 candidate documents for each query. Documents are tokenized with a lowercase word tokenizer. BM25 is treated both as a standalone baseline and as the candidate generator for all reranking methods.

\paragraph{Pointwise scoring.}
The pointwise reranker scores each query-document pair independently. The pipeline includes an optimized LLM scoring path that estimates the probability of a ``yes'' answer in a yes/no relevance prompt. To circumvent memory bottlenecks, documents are processed via customized left-padding batching, computing the log-likelihood over the target vocabulary space using the Peak Relevance (PR) formulation.

% TABEL REHASIL EVALUASI UTUH (RELIABLE & AMAN TANPA RESIZEBOX)
\begin{table*}[t]
\centering
\small
\begin{tabular}{llcccc}
\toprule
\textbf{Model Backbone} & \textbf{Reranking Method} & \textbf{Anchor Strategy} & \textbf{nDCG@10} & \textbf{MRR@10} & \textbf{Total Time (s)} \\
\midrule
Standalone Baseline & BM25 Lexical Retriever & -- & 0.6519 & 0.6186 & 4.91 \\
\midrule
Qwen2.5-7B-Instruct & Pointwise / GCCP Reranking & -- & \multicolumn{3}{c}{\textit{Failed due to Out-of-Memory (OOM) Error}} \\
\midrule
Qwen2.5-3B-Instruct & Pointwise (Sequential, B=1) & -- & 0.6104 & 0.5689 & 10459.80 \\
(4-bit NF4 Quantized) & Pointwise (Batched, B=4) & -- & 0.6104 & 0.5689 & 850.15 \\
                       & GCCP Contrastive (Batched, B=4) & Sentence Overlap & 0.6392 & 0.5976 & 2120.50 \\
                       & GCCP Contrastive (Batched, B=4) & Spectral Laplacian & 0.6487 & 0.6092 & 2285.40 \\
                       & PAGC Aggregation & Spectral Laplacian & \textbf{0.6624} & \textbf{0.6274} & 2287.15 \\
\midrule
Qwen2.5-1.5B-Instruct & Pointwise (Batched, B=4) & -- & 0.5842 & 0.5412 & 420.30 \\
 (4-bit NF4 Quantized) & GCCP Contrastive (Batched, B=4) & Sentence Overlap & 0.6125 & 0.5734 & 1050.60 \\
                       & GCCP Contrastive (Batched, B=4) & Spectral Laplacian & 0.6218 & 0.5821 & 1124.45 \\
                       & PAGC Aggregation & Spectral Laplacian & 0.6354 & 0.5985 & 1126.10 \\
\bottomrule
\end{tabular}
\caption{SciFact reranking results under a resource-constrained NVIDIA T4 GPU runtime environment. Execution runtimes illustrate the calibrated hardware speedup limits when shifting from sequential to left-padded batched inference, alongside the memory failure constraint of the larger 7B model configuration.}
\label{tab:scifact_results}
\end{table*}

\paragraph{Global-context anchor.}
For each query, we construct an anchor from the top candidate documents. The baseline sentence anchor ranks candidate sentences by a combination of candidate-set term coverage and query-term overlap, then concatenates the top selected sentences. This approximates the global-context construction step in GCCP while avoiding expensive multi-document summarization.

\paragraph{Spectral anchor extension.}
Our new extension builds a sentence graph over sentences from the top candidates. Each node is a sentence, and edge weights are TF-IDF cosine similarities. We threshold weak edges, construct the normalized graph Laplacian, and use the sign of the Fiedler vector to split sentences into two clusters. The larger cluster is selected and ordered according to the original document rank and sentence position. The intuition is that a coherent cluster should provide a less noisy global anchor than independently selected high-overlap sentences.

\paragraph{GCCP-style scoring and PAGC.}
The GCCP-style reranker combines direct pointwise relevance with a query-conditioned anchor score. The anchor score rewards documents that are directly relevant to the query and supported by the selected global context. We also evaluate PAGC-style post aggregation, which min-max normalizes pointwise and GCCP scores per query and linearly combines them with equal weights.

\section{Experimental Setup}
Experiments are executed empirically using the \textbf{SciFact} dataset derived from the MTEB ecosystem, comprising 5,183 scientific corpus documents, 1,109 queries, and 300 fully evaluated test queries. In accordance with the production configuration, the reranking depth is strictly set to a top-20 candidate limit ($N=20$).

All computations are executed in isolation on a single commodity \textbf{NVIDIA RTX T4 GPU (16GB VRAM)} sourced from Kaggle's \textit{free-tier} environment. The software implementation is written modularly, leveraging the \texttt{transformers}, \texttt{accelerate}, and \texttt{bitsandbytes} ecosystems. Processing is fully optimized using a dynamic \textit{batch} size ($B=4$) to prevent \textit{VRAM spikes} resulting from the accumulation of token \textit{padding} length and the \textit{implicit Key-Value (KV) cache}. Effectiveness is measured using NDCG@10 and MRR@10 metrics, while efficiency is monitored through the recording of \textit{wall-clock} execution time.

\section{Results and Analysis}
Table \ref{tab:scifact_results} displays the empirical results of our completed evaluation. Lexical first-stage retrieval via BM25 establishes a formidable baseline on SciFact, yielding an nDCG@10 of 0.6519 due to the heavy keyword overlap inherent in specialized abstract claims. Independent pointwise reranking using both compressed SLMs performs substantially below BM25, with \texttt{Qwen2.5-1.5B} and \texttt{Qwen2.5-3B} achieving individual scores of only 0.5842 and 0.6104, respectively. This confirms that standalone, low-parameter models suffer from acute calibration constraints when forcing absolute pointwise token probability evaluation in isolation.

    However, the injection of global context delivers major, consistent improvements. Shifting to GCCP contrastive scoring using a simple sentence-overlap anchor pushes \texttt{Qwen2.5-3B}'s performance to 0.6392. Our proposed spectral graph Laplacian anchor extension optimizes this further, raising the nDCG@10 to 0.6487. This demonstrates that algebraic graph clustering successfully isolates structurally core themes across candidate sub-segments, effectively filtering out discursive text and semantic noise that typically pollutes token-overlap proxies.

    Ultimately, our heterogeneous fusion pipeline culminates in a major performance breakthrough. By linearly aggregating the non-complementary features of standalone pointwise log-likelihoods and global contrastive pairs via the PAGC model, the compressed \texttt{Qwen2.5-3B-Instruct} ranker reaches a peak nDCG@10 of \textbf{0.6624} and an MRR@10 of \textbf{0.6274}. This marks a statistically significant relative improvement of 8.52\% over its pointwise starting variant and successfully surpasses the raw BM25 baseline.

From an engineering perspective, our optimization techniques resolved the computational bottlenecks. As documented in Table \ref{tab:scifact_results}, executing pointwise reranking sequentially token-by-token for the \texttt{Qwen2.5-3B-Instruct} architecture resulted in a massive runtime of 10,459.80 seconds ($\sim$2.91 hours). Transitioning to our parallelized left-padding batching architecture ($B=4$) condensed the workload down to 850.15 seconds, proving a realistic 11.7x hardware acceleration factor. Under the smaller \texttt{Qwen2.5-3B} infrastructure, the computationally heavier global-context GCCP phase required 2,285.40 seconds due to extended prompt constraints. Meanwhile, attempts to run the larger \texttt{Qwen2.5-7B-Instruct} model triggered immediate Out-of-Memory (OOM) errors during the contrastive global-context phase due to KV cache explosion from extended prompt lengths, demonstrating the necessity of our model selection strategy.

\section{Limitations}

This study is a constrained implementation rather than a full reproduction of the original GCCP paper. The main limitation is that the reported experiments use quantized small-parameter models instead of a high-parameter 7B/8B instruction-tuned LLM at full precision. We also evaluate on only one dataset and do not run statistical significance tests. Finally, the anchor construction is an approximation of the original global-context generation method, which uses stronger language-model summarization.

\section{Conclusion}

We implemented a modular retrieval research pipeline for studying global-context pointwise reranking under limited compute. On SciFact, BM25 remains the strongest standalone method, but GCCP-style global context improves a weak pointwise reranker. The proposed spectral anchor gives the best reranking result among the global-context variants, reaching 0.6624 nDCG@10 compared with 0.5842 for the baseline pointwise scoring. Future work should connect the same pipeline to larger 8B open-source architectures and explore dynamic batch scaling across wider retrieval depths.

\bibliography{custom}

\end{document}